---
title: "EDAwR-Code"
format: html
---

```{r}
library(tidyverse)
dat <- read.csv("sample1000.csv")
```

## The data


```{r}
head(dat)
dim(dat)
```

format: *cases* or *unbinned*

(no count or frequency column)


## Mother's age



```{r}
ggplot(dat, aes(mager)) +
  geom_bar(color = "blue", fill = "cornflowerblue") +
  xlab("Mother's age in years")
```

unbinned data --> `geom_bar()`

## Father's age



```{r}
ggplot(dat, aes(fagecomb)) +
  geom_bar(color = "blue", fill = "cornflowerblue") +
  xlab("Fathers's age in years")
```


## Mother's age (father's age is missing)



```{r}
dat |> filter(fagecomb == 99) |> 
ggplot(aes(mager)) +
  geom_bar(color = "blue", fill = "cornflowerblue") +
  xlab("Mother's age in years")
```

## Mother's age by father's age reporting



Create a new variable to indicate missing status

```{r}
dat$father_age_missing <- ifelse(dat$fagecomb == 99, "missing", "reported")
ggplot(dat, aes(mager)) +
  geom_bar(color = "blue", fill = "cornflowerblue") +
  facet_wrap(~father_age_missing)
  
```

## Mother's age by father's age reporting



```{r}
dat |>
  ggplot(aes(mager, fill = father_age_missing)) +
  geom_bar(position = "fill") +
  labs(x = "mother's age", y = "proportion") +
  theme(legend.position = "bottom")
```


## Father's age vs. mother's age


Bin data first. (Not practical to make a scatterplot with almost 4 million data points.)

```{r}
df_count <- dat |> group_by(mager, fagecomb) |> count()
head(df_count)
dim(df_count)
```

## Father's age vs. mother's age


```{r}
g <- df_count |> 
  filter(fagecomb != 99) |> 
  ggplot(aes(mager, fagecomb, fill = n)) + 
  geom_tile() + 
  scale_fill_viridis_c() +
  coord_fixed()
g
```


## Father's age vs. mother's age


```{r}
#| fig-width: 6
#| fig-height: 6
plotly::ggplotly(g)
```

## Most popular ages


```{r}
df_count |> group_by(mager) |> summarize(count = sum(n)) |> 
  slice_max(count, n = 4)
df_count |> group_by(fagecomb) |> summarize(count = sum(n)) |> 
  slice_max(count, n = 4)
```

## Mother's education



```{r}
ggplot(dat, aes(meduc)) +
  geom_bar()
```

::: {.pt7}
(x-axis is incorrect)
:::

## Mother's education

```{r}
ggplot(dat, aes(factor(meduc))) + geom_bar()
```

## Mother's education

```{r}
dat$educ <- fct_recode(factor(dat$meduc), "8th" = "1", "9-12" = "2",
                       "HS" = "3", "Some college" = "4", "Assoc"= "5",
                       "Bachelor" = "6", "Master" = "7", "Doctor" = "8",
                       "Unknown" = "9")
ggplot(dat, aes(educ)) +
  geom_bar() +
  coord_flip()
```


## Mother's age and education



Convert continuous variable to categorical:

```{r}
dat$m_age <- cut(dat$mager, breaks = seq(10, 50, 5))
head(dat)
```

## Mother's age and education



```{r}
#| fig-width: 6
ggplot(dat, aes(educ, fill = fct_rev(m_age))) +
  geom_bar(position = "fill") +
  coord_flip() +
  scale_fill_brewer(direction = -1, breaks = levels(dat$m_age)) +
  guides(fill=guide_legend(title="Mother's Age")) +
  theme_bw(12) +
  theme(legend.position = "bottom")
```


## Mother's age and education



```{r}
ggplot(dat, aes(educ)) +
  geom_bar() +
  coord_flip() +
  facet_wrap(~m_age) +
  scale_y_continuous(labels = ~.x/100000) +
  ylab("Births (in 100,000s)") +
  theme_bw(12)
```

## Mother's age and education


```{r}
ggplot(dat, aes(m_age)) +
  geom_bar() +
  coord_flip() +
  facet_wrap(~educ) +
  scale_y_continuous(labels = ~.x/100000) +
  ylab("Births (in 100,000s)") +
  theme_bw(12)
```

